---
title: "p8105_hw3_yx2638"
author: "Yifei Xu"
date: "2022-10-15"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8,
	fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

```{r}
# load the "Instacart" data from the "p8105.datasets"
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

The dataset `instacart` contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.


```{r}
# summarize the number of items ordered from aisle
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```
In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
# plot the number of items ordered in each aisle
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) 

```

```{r}
# Show the three most popular items in aisles `baking ingredients`, ` dog food care`, and `packaged vegetables fruits`,and include the number of times each item is ordered 
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

```{r}
# show the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

# Problem 2

#### Read in the data 

```{r}
accel = read_csv("data/accel_data.csv") 
```

#### Tidy and wrangle the data

```{r}
accel_tidy = accel %>%
        janitor::clean_names() %>%
        pivot_longer(activity_1:activity_1440, names_to = "minute", values_to = "activity_counts", names_prefix =  "activity_") %>%
        mutate(day = day %>% 
                       fct_relevel("Monday", "Tuesday", "Wednesday","Thursday","Friday", "Saturday", "Sunday")) %>% 
        mutate(weekday_vs_weekend = ifelse(day %in% c("Saturday", "Sunday"), "weekend", "weekday")) %>%
        mutate(minute = as.integer(minute)) 
```

The resulting dataset contains `r nrow(accel_tidy)` observations. `r ncol(accel_tidy)` key variables include `r names(accel_tidy)`. Each observation represent a male's activity count per minute of each day during five weeks.


#### Create a table to show the total activity counts

```{r}
# aggregate across minutes to create a total activity variable for each day, and create a table showing these totals
accel_tidy %>%
        group_by(week, day) %>%
        summarize(total_activity_counts = sum(activity_counts)) %>%
        knitr::kable()
```

We cannot observe apparent trend from the table. However, it's obvious that the total activity accounts on Saturdays in week 4 and week 5 is far lower than any other days.

#### Make a single-panel plot that shows the 24-hour activity time courses for each day

```{r}
accel_tidy %>%
        group_by(day, minute) %>%
        ggplot(aes(x = minute, y = activity_counts)) +
        geom_point(aes(color = day), alpha = 0.5) +
        scale_color_brewer(palette = "Set1") +
        geom_smooth() + 
        scale_x_continuous(
                breaks = c(0, 180, 360, 540, 720, 900, 1080, 1260, 1440),
                labels = c("12am", "3am", "6am", "9am", "12pm", "3pm", "6pm", "9pm", "12am")) + 
        labs(x = "Time", 
             y = "Activity Counts", 
             title = "24-hour Activity Time Courses for Each Day")
```

Within a day, we can observe that the activity count of this male is very low between 12am and 3am, and then begins to increase. Around 9-12w wwwam, the activity count reaches the first peak, after then it remains with little fluctuation. The second peak occurs around 9pm, after which the activity account begins to decline.

Within weekday, his activity count is relatively high around 7am on some Thursdays and around 9am on some Fridays. As for the weekend, the activity count is high in the afternoon, especially around 4-5pm on Saturday and around 12pm on Sunday. Between 8-10pm, the count is usually high across many days, especially on Friday.

From this plot, we can roughly identify his daily schedule (such as the time of wake-up and sleep).

# Problem 3

#### Read in the data

```{r}
data("ny_noaa")

ny_noaa = 
  ny_noaa %>% 
  as_tibble(ny_noaa)
```


